---
title: "homework2final"
author: "NinaE"
date: "September 27, 2015"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---


Loading data:

```{r}
library(gapminder)
str(gapminder)
```

I am going to small test the data:

```{r}
mode(gapminder)
class(gapminder)
```

Gapfinder is a data.frame and it's in "list" mode

```{r}
head(gapminder)
tail(gapminder)
gapminder[sample(nrow(gapminder), 10), ]
View(gapminder)
names(gapminder)
ncol(gapminder)
nrow(gapminder)
dim(gapminder)
```

It has 6 variables, 6 columns and 1704 rows

Variable flavors:
county, continent: factor variables
Year, lifeExp, pop and gdpPercap: numerical variables

Exploring variables:

```{r}
class(gapminder$continent)
summary(gapminder$continent)
table(gapminder$contintent)
levels(gapminder$continent)
nlevels(gapminder$continent)
unclass(gapminder$continent)
class(gapminder$lifeExp)
summary(gapminder$lifeExp)
table(gapminder$lifeExp)
levels(gapminder$lifeExp)
nlevels(gapminder$lifeExp)
unclass(gapminder$lifeExp)

plot(lifeExp ~ continent, gapminder)
plot(lifeExp ~ year, gapminder)
plot(gapminder$continent, gapminder$lifeExp)
plot(lifeExp ~ log(gdpPercap), gapminder)
plot(lifeExp ~ country, gapminder)
plot(lifeExp ~ continent, gapminder)
```

Exploring plot types:

barplots

```{r}
library(ggplot2)
barplot(table(gapminder$country))
barplot(table(gapminder$continent))
ggplot(gapminder, aes(x = continent)) + geom_bar()
p <- ggplot(gapminder, aes(x = reorder(continent, continent, length)))
p + geom_bar()
p + geom_bar(width = 0.05)

```

scatterplots:

```{r}
p <- ggplot(gapminder, aes(x = pop, y = lifeExp))
p + geom_point()

ggplot(gapminder, aes(x = log10(pop), y = lifeExp)) + geom_point()
p + geom_point() + scale_x_log10()
p <- p + scale_x_log10()
p + geom_point(aes(color = continent))
ggplot(gapminder, aes(x = pop, y = lifeExp, color = continent)) + geom_point() + scale_x_log10()
p + geom_point(alpha = (1/3), size = 3)
p + geom_point() + geom_smooth()
p + geom_point() + geom_smooth(lwd = 3, se = FALSE, method = "lm")
p + aes(color = continent) + geom_point() + geom_smooth(lwd = 3, se = FALSE)
p + aes(color = continent) + geom_point() + geom_smooth(lwd = 3, se = FALSE, method = "lm")
p + geom_point(alpha = (1/3), size = 3) + facet_wrap(~ continent)
geom_smooth(lwd = 2, se = FALSE)
geom_smooth(lwd = 2, se = FALSE, method = "lm")
```

stripplots

```{r}
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_point()
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_jitter()
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```

boxplots

```{r}
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_boxplot()
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_boxplot(outlier.colour = "blue") + geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4) + stat_summary(fun.y = median, colour = "red", geom = "point", size = 2)
ggplot(gapminder, aes(reorder(x = continent, lifeExp), y = lifeExp)) +geom_jitter(position = position_jitter(width = 0.1), alpha = 1/4) + stat_summary(fun.y = median, colour = "red", geom = "point", size = 2)
```

histogram

```{r}
ggplot(gapminder, aes(x = lifeExp, fill = continent)) + geom_histogram()
ggplot(gapminder, aes(x = lifeExp, color = continent)) + geom_density()
ggplot(gapminder, aes(x = lifeExp, fill = continent)) + geom_density(alpha = 0.2)
```


I used the different ggplot tutorials supplied by Jenny. 